---
title: "Diversidade Genetica"
author: "LucianoRogerio"
date: "2021-11-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Shannon Weaver Index estimation for each core collection

### Shannon Weaver Index Quantitative traits

```{r Preparing the Nuclear Core Collection data}
suppressMessages(library(tidyverse)); suppressMessages(library(data.table)); suppressMessages(library(reactable))
source(here::here("code", "Shannon-Weaver.R"))
Alldataset <- readRDS(here::here("output", "AllDataCCCaroline.RDS"))

QualityTrait <- colnames(Alldataset)[sapply(X = Alldataset, class) == "character"] %>% 
  setdiff(., c("Acessos", "Data"))


AlldataSetQuant <- Alldataset %>% dplyr::select(-all_of(QualityTrait))
AlldataSetQuali <- Alldataset %>% dplyr::select(Acessos, QualityTrait, Data) %>% suppressMessages()

QuantTrait <- data.frame(Trait = colnames(AlldataSetQuant)[2:17],
                         Min = apply(AlldataSetQuant[2:17], FUN = min, na.rm = TRUE, MARGIN = 2),
                         Max = apply(AlldataSetQuant[2:17], FUN = max, na.rm = TRUE, MARGIN = 2))
DataMethod <- unique(data.frame(Data = AlldataSetQuant$Data))
```

```{r Estimating the Shannon-Weaver Index for the quantitative traits}
QuantSHs <- NULL
for(i in 1:nrow(DataMethod)){
  for(trait in QuantTrait$Trait){
    QuantSH <- data.frame(Data = DataMethod$Data[i],
                          Trait = trait,
                          SH = Shannon.Weaver.QT(AlldataSetQuant[(AlldataSetQuant$Data%in%DataMethod$Data[i]),trait],
                                                 Min = QuantTrait[QuantTrait$Trait == trait, "Min"],
                                                 Max = QuantTrait[QuantTrait$Trait == trait, "Max"],
                                                 formula = 2))
    QuantSHs <- rbind(QuantSHs, QuantSH)
  }
}

Table1SH <- matrix(QuantSHs$SH, byrow = F, ncol = 4)
rownames(Table1SH) <- unique(QuantSHs$Trait)
colnames(Table1SH) <- unique(paste(QuantSHs$Data, sep = " - "))
colnames(Table1SH)[1] <- "BAG"

write.table(Table1SH, file = here::here("output", "Shannon-WeaverCCQuantIndex.csv"),
            sep = ",", dec = ".", quote = FALSE)
```


#### **Table 1.** Shannon-Weaver indexes for Cassava quantitative traits of the Bank of germplasm and 12 nuclear core collections.
```{r Table 1, echo = FALSE}
Table1SH %>% reactable(defaultPageSize = 16, columns = list(
  BAG = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Gen/Phen' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  Phen = colDef(format = colFormat(digits = 3, locales = "en-US")),
  Gen = colDef(format = colFormat(digits = 3, locales = "en-US"))))
```

### Shannon Weaver Index Qualitative traits

```{r Estimating the Shannon-Weaver for the Qualitative traits}

Trait <- c("CorFolhaApical" = 4, "CorPeciolo" = 6, "CorFolhaDesenv" = 5,
           "CorNerv" = 4, "CorCortexCaule" = 4, "CorExternaCaule" = 7,
           "CorEpidCaule" = 4, "CorRamosTerm" = 3, "CorExternaRzs" = 4,
           "CorCortexRzs" = 4, "CorPolpaRzs" = 5, "NLobulos" = 5,
           "Pubescencia" = 2, "FormaLobuloCentral" = 9, "PosicaoPeciolo" = 4,
           "Sinuosidade" = 2, "ComprEstip" = 2, "HabCrescCaule" = 2,
           "TipoPlan" = 4, "MargEstip" = 2, "HabRamif" = 4, "NiveisRam" = 4, 
           "ProemCicatrizesFolhas" = 2, "FormaRzs" = 4, "TxtEpidermeRzs" = 2,
           "PresPedunculoRzs" = 3, "PosicaoRzs" = 3, "Flowering" = 2,
           "DestqPelicRz" = 2, "DestqCortexRz" = 3, "ConstrRzs" = 3,
           "AnguloRamif" = 22)

QualTrait <- data.frame(Trait = names(Trait),
                        NClas = Trait,
                        row.names = NULL)

QualSHs <- NULL
for(i in 1:nrow(DataMethod)){
  for(trait in QualTrait$Trait) {
    Data <- AlldataSetQuali[(AlldataSetQuali$Data%in%DataMethod$Data[i]),trait]
    ShannonWeaver <- Shannon.Weaver.QL(Data,
                                       nclass = QualTrait$NClas[QualTrait$Trait%in%trait])
    QualSH <- data.frame(Data = DataMethod$Data[i],
                         Trait = trait,
                         SH = ShannonWeaver)
    QualSHs <- rbind(QualSHs, QualSH)
  }
}

Table2SH <- matrix(QualSHs$SH, byrow = F, ncol = 4)
rownames(Table2SH) <- unique(QualSHs$Trait)
colnames(Table2SH) <- unique(paste(QualSHs$Data, sep = " - "))
colnames(Table2SH)[1] <- "BAG"

write.table(Table2SH, file = here::here("output", "Shannon-WeaverCCQualIndex.csv"),
            sep = ",", dec = ".", quote = FALSE)

SHInd <- as.data.frame(rbind(Table1SH, Table2SH))
```


#### **Table 2.** Shannon-Weaver indexes for Cassava qualitative traits of the Bank of germplasm and 12 nuclear core collections.
```{r Table 2, echo = FALSE}
Table2SH %>% reactable(defaultPageSize = 15, columns = list(
  BAG = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Gen/Phen' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Phen' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Gen' = colDef(format = colFormat(digits = 3, locales = "en-US"))))
```

## Improvement in Shannon-Weaver index in Core Collection
```{r Selection of the most diverse Nuclear Collection estimated with Phenotypic data}
SHBAG <- SHInd[, "BAG"]
SHIndCC <- SHInd[ , !colnames(SHInd)%like%"BAG"]

SHRelCC <- (SHIndCC - SHBAG)*100/SHBAG

write.table(SHRelCC, file = here::here("output", "GanhoRelativoShannonWeaver.csv"),
            quote = F, sep = ",")

SHSelCC <- data.frame(Method = colnames(SHRelCC),
                        SHMeanImprov = colMeans(SHRelCC, na.rm = TRUE),
                        SHsd = apply(SHRelCC, FUN = sd, na.rm = TRUE, MARGIN = 2),
                        row.names = NULL)
```

#### **Table 3.** Mean improvement of the Shannon-Weaver diversity index in % of nuclear core collections estimated of Phenotypic data compared to the Bank of Germplasm
```{r Table 5, echo = FALSE}
reactable(SHSelCC, columns = list(
  SHMeanImprov = colDef(format = colFormat(digits = 3, locales = "en-US")),
  SHsd = colDef(format = colFormat(digits = 3, locales = "en-US"))))
```

The best phenotypic Core Collection selected was CH.


## Mean and variance of each Nuclear Collection

```{r Mean and Variance of Nuclear Collections}
AlldataSetQuant %>% group_by(Data) %>%
  summarise(ComprFilotaxiaMean = mean(ComprFilotaxia, na.rm = T),
            ComprFilotaxiaVar = var(ComprFilotaxia, na.rm = T),
            ComprLobuloMean = mean(ComprLobulo, na.rm = T),
            ComprLobuloVar = var(ComprLobulo, na.rm = T),
            LargLobuloMean = mean(LargLobulo, na.rm = T),
            LargLobuloVar = var(LargLobulo, na.rm = T),
            RelComprLarMean = mean(RelComprLar, na.rm = T),
            RelComprLaraVar = var(RelComprLar, na.rm = T),
            ComprPecioloMean = mean(ComprPeciolo, na.rm = T),
            ComprPecioloVar = var(ComprPeciolo, na.rm = T),
            EspEntreCascaMean = mean(EspEntreCasca, na.rm = T),
            EspEntreCascaVar = var(EspEntreCasca, na.rm = T),
            ComprMedRzsMean = mean(ComprMedRzs, na.rm = T),
            ComprMedRzsVar = var(ComprMedRzs, na.rm = T),
            DiamMedRzsMean = mean(DiamMedRzs, na.rm = T),
            DiamMedRzsVar = var(DiamMedRzs, na.rm = T),
            HCNPicMean = mean(HCNPic, na.rm = T),
            HCNPicVar = var(HCNPic, na.rm = T),
            NRMean = mean(NR, na.rm = T),
            NRVar = var(NR, na.rm = T),
            PAMean = mean(PA, na.rm = T),
            PAVar = var(PA, na.rm = T),
            DRYMean = mean(DRY, na.rm = T),
            DRYVar = var(DRY, na.rm = T),
            DMCsgMean = mean(DMCsg, na.rm = T),
            DMCsgVar = var(DMCsg, na.rm = T),
            PTRMean = mean(PTR, na.rm = T),
            PTRVar = var(PTR, na.rm = T),
            PPAMean = mean(PPA, na.rm = T),
            PPAVar = var(PPA, na.rm = T),
            ICMean = mean(IC, na.rm = T),
            ICVar = var(IC, na.rm = T),
            APMean = mean(AP, na.rm = T),
            APVar = var(AP, na.rm = T)) -> NCMeanVar
write.table(NCMeanVar, file = here::here("output", "MeansVarNC.csv"),
            quote = F, row.names = F, sep = ",")
```

#### **Table 4.** Means and variances for Nuclear Collections of quantitative traits 
```{r Table 3, echo = FALSE}
library(reactable)
NCMeanVar %>%
  reactable::reactable(defaultColDef = colDef(format = colFormat(digits = 2,
                                                                 locales = "en-us")))
```


## Grouping of Nuclear Collection methods by Kappa index

```{r kappa Index estimation}
BAGClones <- Alldataset$Acessos[Alldataset$Data == "BAG"]

CCMethods <- unique(Alldataset$Data) %>% setdiff("BAG")

SelCC <- tibble(Clones = BAGClones)

for(i in CCMethods){
  Common <- BAGClones %in% Alldataset$Acessos[Alldataset$Data == i] %>%
    as.numeric %>% as.data.frame
  colnames(Common) <- i
  SelCC <- cbind(SelCC, Common)
}
row.names(SelCC) <- SelCC$Clones
SelCC$Clones <- NULL

CoincObs <- matrix(data = NA, nrow = ncol(SelCC), ncol = ncol(SelCC))
colnames(CoincObs) <- rownames(CoincObs) <- colnames(SelCC)

for(i in colnames(SelCC)){
  for(j in colnames(SelCC)){
    if(i == j){ CoincObs[i, j] <- 1} else {
      CoincObs[i, j] <- sum(SelCC[, i] == SelCC[, j])/nrow(SelCC)
    }
  }
}

CoincEsp <- matrix(data = (sum(SelCC[,1])/nrow(SelCC))**2 + (1-(sum(SelCC[,1])/nrow(SelCC)))**2,
                   nrow = ncol(SelCC), ncol = ncol(SelCC))
colnames(CoincEsp) <- rownames(CoincEsp) <- colnames(SelCC)

Kappa <- round((CoincObs - CoincEsp)/(1 - CoincEsp), digits = 4)
write.table(Kappa, file = here::here("output", "KappaIndex.csv"), sep = ",",
            quote = F)
```

#### **Table 5.** Kappa Index for coincidence of the clone selection for the 12 nuclear collections.
```{r Table 4, echo = FALSE}
library(reactable)
Kappa %>% reactable(defaultPageSize = 15)
```

#### **Fig 1.** Kappa index plot of the coincidence of selected clones by the 12 methods of nuclear collection.
```{r Fig 1, echo = FALSE}
library(corrplot)
corrplot::corrplot(Kappa,cl.pos = "n", tl.col = c("brown", "red", "green"))
```

### Venn diagram for each Nuclear Collection Group

```{r Venn diagram}
library(ggvenn)
VennDPNC <- list(Phen = rownames(SelCC)[SelCC$Phen == 1],
                 Gen = rownames(SelCC)[SelCC$`Gen` == 1],
                 'Gen/Phen' = rownames(SelCC)[SelCC$`Gen/Phen` == 1])
VD1 <- ggvenn(VennDPNC, fill_color = c("red", "green", "brown"),
              fill_alpha = 0.6, set_name_size = 5)

sink(file = here::here("output", "ClonesSelCC.txt"))
VennDPNC
sink()
```



#### **Fig 2.** Venn diagram of Nuclear Collections
```{r Fig 2, fig.height = 12, echo = FALSE}
VD1
```


## Comparison of the diversity Between BAG and the Gen/Phen Core Collection

```{r PCA analysis of the Gower Distance from the Bank of Germplasm accessions}
suppressMessages(library(StatMatch)); library(here)
suppressMessages(library(tidyverse))

DataCar <- readRDS(file = here::here("output", "DataSelPreparedCar.rds")) %>% 
  mutate(ComprFilotaxia = as.character(ComprFilotaxia),
         ComprEstip = as.character(ComprEstip))

DistCar <- gower.dist(data.x = DataCar)
row.names(DistCar) <- colnames(DistCar) <- rownames(DataCar)
PCA <- prcomp(DistCar)

Perc <- 100*PCA$sdev^2/sum(PCA$sdev^2)
PercAc <- as.vector(rep(NA, times = length(Perc)))
for(i in 1:length(Perc)) {
  PercAc[i] <- sum(Perc[1:i])
  names(PercAc)[i] <- i
}
```

#### **Fig 3.** Accumulated Variance explained by PCA
```{r Fig 3, echo = FALSE}
barplot(PercAc[1:15], main = "Variance explained by PCA",
        ylab = "Cumulative variance (%)", xlab = "Number of retained PCs",
        col = c("gray"), ylim = c(0, 100))
```



```{r Creating columns for the plot legend}
PointPCA1 <- as.data.frame(PCA$x[,1:5]) %>% dplyr::mutate(Collection = "BAG")

Alldataset <- readRDS(here::here("output", "AllDataCCCaroline.RDS"))

PointPCA1 %<>% bind_rows(., PointPCA1[Alldataset %>%
                                       filter(Data == "Gen/Phen") %>% .$Acessos,] %>% 
                           dplyr::mutate(Collection = "Gen/Phen")) %>%
  bind_rows(., PointPCA1[Alldataset %>%
                                       filter(Data == "Gen") %>% .$Acessos,] %>% 
                           dplyr::mutate(Collection = "Gen")) %>%
  bind_rows(., PointPCA1[Alldataset %>%
                                       filter(Data == "Phen") %>% .$Acessos,] %>% 
                           dplyr::mutate(Collection = "Phen")) %>% 
  dplyr::mutate(Collection = factor(Collection, levels = c("BAG", "Phen", "Gen", "Gen/Phen")))
```

```{r PCA ggplots}
PC12 <- ggplot(data = PointPCA1, aes(x = PC1, y = PC2, color = Collection, fill = Collection)) +
  geom_point() +
  geom_rug(col = "steelblue", alpha = 0.2, size = 1.5) +
  scale_shape_manual(values = c(16, 16, 16, 1)) +
  scale_color_manual(values = c("blue", "green", "red", "brown")) +
  stat_ellipse(geom="polygon", 
               alpha = 0.2, 
               show.legend = FALSE, 
               level = 0.95) +
  xlab(paste0("PC1 - ", round(Perc[1], digits = 2), "%")) +
  ylab(paste0("PC2 - ", round(Perc[2], digits = 2), "%"))

PC34 <- ggplot(data = PointPCA1, aes(x = PC3, y = PC4, color = Collection, fill = Collection)) +
  geom_point() +
  geom_rug(col = "steelblue", alpha = 0.2, size = 1.5) +
  scale_shape_manual(values = c(16, 16, 16, 1)) +
  scale_color_manual(values = c("blue", "green", "red", "brown")) +
  stat_ellipse(geom="polygon", 
               alpha = 0.2, 
               show.legend = FALSE, 
               level = 0.95) +
  xlab(paste0("PC3 - ", round(Perc[3], digits = 2), "%")) +
  ylab(paste0("PC4 - ", round(Perc[4], digits = 2), "%"))
```

#### **Fig 4.** Análise de Componentes Principais da Matriz de Distância de Gower dos acessos do Banco ativo de germoplasma da EMBRAPA Selecionados pelas Coleções Nucleares
```{r Fig 4, echo = FALSE}
library(ggpubr)
ggarrange(PC12, PC34, common.legend = T, legend = "right", ncol = 1, labels = "AUTO")
```



Back - [Análises Descritivas](AnDis.html)

[Home](index.html)
